Formalizing Stack Safety as a Security Property

### 2021-07-08 submission3/main.tex
pass through 7

\section{Threat Model, Machines, and Enforcement }
\label{sec:setup}

@@

We now formalize the model of enforcement that we introduced in \Cref{sec:threat}.
A {\em policy} consists of a set of policy
states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
it is undefined on input configurations that correspond to a policy fault. The policy
can see the machine state, but cannot modify it. The policy also provides a
well-formedness predicate, \(\WF \subseteq \MACHS \times \POLS\), describing
initial states that are properly configured.
%
%A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
%\Cref{sec:enforcement}.---APT: We just said this.

@@

The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
A call annotation tells us to push the return target onto the return stack and
seal all components with the depth of the previous return stack. If we aren't making a call,
we identify a return by looking ahead and checking if we're about to step to a state
that matches the topmost return target. If neither is true, the context remains unchanged.
Recall that a ``code map'' \(\codemap\) is a partial function, and we can use its domain
as a predicate on addresses.

@@

\definition
{\em Observational integrity} means that for any \(\MPCT,\brk \in \MPCT^*\)
where \(\head(\MPCT) = \mach,\pol,\context\) and \(\last(\MPCT) = \mach',\pol',\context'\),
the following holds. Let \(\components\) be the set of protected components that changed during
the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
Then for any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), and its induced trace
\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).

@@

\definition
A system enjoys {\em observational confidentiality} if for all
\(\MPCT,\brk \in \MPCT^*\), with \(\head(\MPCT) = \mach,\pol,\context\) and
\(\components = \secret(\context)\), the following holds.
%
For any \(\nach\) such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \brk(\MPCT)\),
\(\MPCT\) and \(\NPCT\) respect three conditions:

@@

\paragraph{Formalization of variable passing.}
We add extra information to our call
annotations to reflect that they may now pass variables on the stack, and that the locations
of those variables may be dynamic (as in the case of pass-by-reference).
A call annotation now includes a predicate relating machine states to the components
that should be passed, expressed as offsets from the stack pointer.
%
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P \subseteq \WORDS)\]

We extend the set of domains to include passed components,
which are labeled with the range of depths of the caller that has passed them.
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
\times (\listT (\target)) \]
The initial context again maps all components to \(\unsealed\).
We straightforwardly describe how the sealing convention and passing predicate combine to
update a domain map at a call from depth \(\depth\).
\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
\begin{cases}
  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and }
                                 \component - (\mach ~ \rsp) \not \in P \\
  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and }
                                 \component - (\mach ~ \rsp) \in P \\
  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
                                 \sealcon ~ \mach ~ \component \\
                               & \text{ and } (\mach ~ \rsp) \in P, \text{ for any } \depth' \\
  \domMap~\component & \text{else} \\
\end{cases}\]

@@

\definition Let the protected set of a context, \(\prot_P(\domMap,\rts)\),
be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
\(|\rts| > \depth'\).

\definition Let the \emph{secret set} of a context, written \(\secret_P(\domMap,\rts)\),
be the set of components \(\component\) such that
\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
\(\domMap ~ \component = \passed{\depth,\depth'}\)
where \(|\rts| < \depth\) or \(|\rts| > \depth'\).

\definition
For each of the stepwise, deferred, and observational versions of
stack integrity in the simple model, the equivalent property in the
passing model substitutes the above \(\CONTEXTS\) type and update
function, and \(\prot_P\), for their equivalents in the simple model.

For both the stepwise and observational versions of stack confidentiality in
the original model, the passing model equivalents additionally substitute \(\secret_P\).

@@

Intuitively, a caller expects a callee-saves register to be unchanged on
return: it
may change during the call, as long as it is restored. We similarly expect a callee-saves
register to be secret from the callee, except when it is copied to the stack! Both scenarios
reflect not stepwise but observational stack safety.

We begin by adding information to Call annotations defining which registers will be {\em saved}
and which are {\em unsaved}. The stack pointer and the register containing the return address
should never be saved. We then extend the set of domains with these values.
%
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(S \in \REGS \rightarrow (\unsaved + \saved))\]
%
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \unsaved + \saved\]

\judgmentthree[ Call]
              {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
%              {\(\components = \sealcon ~ \mach\)}
              {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \in \REGS \mapsto S \reg]\)}
              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}

@@

\definition
%
{\bf Observational stack integrity} with callee-saves registers substitutes the above
\(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
%
{\bf Observational stack confidentiality} additionally substitutes \(\secret_S\).

@@

Observational properties are not only for lazy enforcement mechanisms---they reflect a useful
loosening of the stepwise properties. Consider a single register \(\reg\) that is saved.
It is copied to some stack address \(\addr\), violating stepwise confidentiality, and then
overwritten, violating stepwise integrity. Then before the return, it is restored to its
original value. In observational integrity terms, \(\reg\) is unchanged between call and return,
and in observational confidentiality terms, both the original and variant values are unchanged
from their originals, so that \(\reg\) is not in the \(\delta\) set. But \(\addr\) is in the
\(\delta\) set, so confidentiality additionally requires that the value of \(\reg\) does not
leak through the callee-saves mechanism, for instance if the caller allocates but fails to
initialize more memory.

@@

@@

This calls for some explanation. First, we determine which coroutine we are yielding to by
kchecking which stack the stack pointer of the next state is pointing to. Then we can only yield
if the next state is a valid yield target for that stack, and we update the yield target for
the current stack, ensuring that when the system yields back we expect the current coroutine
to continue from where it left off. So, it is possible for a coroutine to claim to yield
but not actually change the active stack, with similar effects to when a callee fails
to return properly. Now we can define both stack safety and coroutine safety under this model.

@@

\definition \(\MPCT^\wedge\) is the \emph{yield set}, defined as
\[\begin{aligned}
\bigcup_{\stackid \in \STACKS} \hspace{-.5em} \{\MPCT, \neg f_\stackid | \MPCT\langle f_\stackid \rangle
\text{ is a reachable segment}\} \hspace{1em} &
\text{where } f_\stackid ~ (\underscore,\underscore,(\underscore,\underscore,\underscore,\stackid')) \triangleq \stackid \not = \stackid' \\
\end{aligned}\]

Here, we are capturing the entire span that the system spends {\em away} from a given
coroutine.

\definition
Let the protected set of a context, \(\prot_Y(\domMap,\rtm,\yts,\stackid)\),
be the set of components \(\component\) such that
\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).

\definition
Let the secret set of a context, \(\secret_Y(\domMap,\rtm,\yts,\stackid)\),
be the set of components \(\component\) such that
\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).

>>>>

\section{Threat Model, Machines, and Enforcement }
\label{sec:setup}
\label{sec:threat}

@@

We now formalize the model of enforcement that we introduced in Section~\ref{sec:threat}.
A {\em policy} consists of a set of policy
states \(\pol \in \POLS\) and a policy step function \((\mach, \pol) \polstep \pol' \in
\MACHS \times \POLS \rightharpoonup \POLS\). This step function is partial;
it is undefined on input configurations that correspond to a policy fault. The policy
can see the machine state, but cannot modify it. The policy also provides a
well-formedness predicate, \(\WF \subseteq \MACHS \times \POLS\), describing
initial states that are properly configured.
%
%A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
%\Cref{sec:enforcement}.---APT: We just said this.

@@

The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
A call annotation tells us to push the return target onto the return stack and
seal all components with the depth of the previous return stack. If we aren't making a call,
we identify a return by looking ahead and checking if we're about to step to a state
that matches the topmost return target. If neither is true, the context remains unchanged.
Recall that a ``code map'' \(\codemap\) is a partial function, and we can use its domain
as a predicate on addresses.\apt{We never explain what $\ANNS$ is for the code maps in
  this section.}

@@

\definition
\textbf{\em Observational integrity} means that for any \(\MPCT,\brk \in \MPCT^*\)
where \(\head(\MPCT) = \mach,\pol,\context\) and \(\last(\MPCT) = \mach',\pol',\context'\),
the following holds. Let \(\components\) be the set of protected components that changed during
the call, \(\components = \prot(\context) \cap \Delta(\mach, \mach')\),
and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
Then for any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), and its induced trace
\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).

@@

\definition
A system enjoys \textbf{\em observational confidentiality} if for all
\(\MPCT,\brk \in \MPCT^*\), with \(\head(\MPCT) = \mach,\pol,\context\) and
\(\components = \secret(\context)\), the following holds.
%
For any \(\nach\) such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \brk(\MPCT)\),
\(\MPCT\) and \(\NPCT\) respect three conditions:

@@

\paragraph{Formalization of variable passing.}
We add extra information to our call
annotations to reflect that they may now pass variables on the stack, and that the locations
of those variables may be dynamic (as in the case of pass-by-reference).
A call annotation now includes a predicate relating machine states to the components
that should be passed, expressed as offsets from the stack pointer.
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(P)~\mbox{\rm where}~P \subseteq \WORDS\]
We extend the set of domains to include passed components,
which are labeled with the range of depths of the caller that has passed them.
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \passed{\depth,\depth'}\]
The definition of context looks the same, but is based on the revised definition of \(\STKDOMS\):
\[(\domMap,\rts) = \context \in \CONTEXTS ::= (\COMPONENTS \rightarrow \STKDOMS)
\times (\listT (\target)) \]
The initial context again maps all components to \(\unsealed\).
We straightforwardly describe how the sealing convention and passing predicate combine to
update a domain map at a call from depth \(\depth\).
\[(\mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ \depth) ~ \component =
\begin{cases}
  \sealed{\depth}              & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and }
                                 \component - (\mach ~ \rsp) \not \in P \\
  \passed{\depth,\depth+1}     & \text{if } \domMap ~ \component = \unsealed \text{ and }
                                 \sealcon ~ \mach ~ \component \text{ and }
                                 \component - (\mach ~ \rsp) \in P \\
  \passed{\depth',\depth+1}    & \text{if } \domMap ~ \component = \passed{\depth',\depth} \text{ and }
                                 \sealcon ~ \mach ~ \component \\
                               & \text{ and } (\mach ~ \rsp) \in P, \text{ for any } \depth' \\
  \domMap~\component & \text{else} \\
\end{cases}\]

@@

\definition The \emph{protected set} of a context, \(\prot_P(\domMap,\rts)\),
is the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\underscore}\)
or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
\(|\rts| > \depth'\).

\definition The \emph{secret set} of a context, written \(\secret_P(\domMap,\rts)\),
is the set of components \(\component\) such that
\(\domMap ~ \component = \sealed{\underscore}\), \(\domMap ~ \component = \unsealed\), or
\(\domMap ~ \component = \passed{\depth,\depth'}\)
where \(|\rts| < \depth\) or \(|\rts| > \depth'\).

\definition
For each of the \textbf{\em stepwise}, \textbf{\em deferred}, and \textbf{\em observational} versions of
\textbf{\em stack integrity} in the simple model, the equivalent property in the
passing model substitutes the above \(\CONTEXTS\) type and update
function, and \(\prot_P\), for their equivalents in the simple model.

For both the \textbf{\em stepwise} and \textbf{\em observational} versions of \textbf{\em stack confidentiality} in
the original model, the passing model equivalents additionally substitute \(\secret_P\).

@@

Intuitively, a caller expects a callee-saves register to be unchanged on
return: it
may change during the call, as long as it is restored. We similarly expect a callee-saves
register to be secret from the callee, except when it is copied to the stack! Both scenarios
reflect observational rather than stepwise stack safety.

\apt{Mention that now for the first time we are putting register components into something other than the outside domain.}
We begin by adding information to Call annotations defining which registers will be {\em saved}
and which are {\em unsaved}. The stack pointer and the register containing the return address
should never be saved. We then extend the set of domains with these values.
%
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \callmap(S) ~ \mbox{\rm{where}} ~ S \in \REGS \rightarrow (\unsaved + \saved)\]
%
\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth} + \unsaved + \saved\]

\judgmentthree[ Call]
              {\(\codemap ~ (\mach ~ \PCname) = \callmap(S)\)}
%              {\(\components = \sealcon ~ \mach\)}
              {\(\components = \{\component | \sealcon ~ \mach ~ \component\}\)}
              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}][\reg \mapsto S(\reg) ~ | ~ r \in \REGS]\)}
              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}

@@

\definition
%
\textbf{\em Observational stack integrity} with callee-saves registers substitutes the above
\(\CONTEXTS\) type and update function, and \(\prot_S\), for its equivalent in the simple model.
%
\textbf{\em Observational stack confidentiality} additionally substitutes \(\secret_S\).

@@

Observational properties are not only for lazy enforcement mechanisms---they reflect a useful
loosening of the stepwise properties. Consider a single register \(\reg\) that is saved.
It is copied to some stack address \(\addr\), violating stepwise confidentiality, and then
overwritten, violating stepwise integrity. Then before the return, it is restored to its
original value. In observational integrity terms, \(\reg\) is unchanged between call and return,
and in observational confidentiality terms, both the original and variant values are unchanged
from their originals, so that \(\reg\) is not in the \(\delta\) set.\apt{?? what's that?}  But \(\addr\) is in the
\(\delta\) set, so confidentiality additionally requires that the value of \(\reg\) does not
leak through the callee-saves mechanism, for instance if the caller allocates but fails to
initialize more memory.

@@

\apt{Should that be $\passed{\depth,\depth'}$ ?.  And why do we need to be precise about $\STKDOMS$ anyhow?
  And shouldn't the first component of $\context$ be of type $\COMPONENTS \rightarrow \TOPDOMS$? }

@@

This calls for some explanation. First, we determine which coroutine we are yielding to by
checking which stack the stack pointer of the next state is pointing to. Then we can only yield
if the next state is a valid yield target for that stack, and we update the yield target for
the current stack, ensuring that when the system yields back we expect the current coroutine
to continue from where it left off. So, it is possible for a coroutine to claim to yield
but not actually change the active stack, with similar effects to when a callee fails
to return properly. Now we can define both stack safety and coroutine safety under this model.

@@

\definition The \emph{yield set} \(\MPCT^\wedge\) of a trace \(\MPCT\) \apt{right?} is
\[\begin{aligned}
\bigcup_{\stackid \in \STACKS} \hspace{-.5em} \{\MPCT, \neg f_\stackid | \MPCT\langle f_\stackid \rangle
\text{ is a reachable segment}\} \hspace{1em} &
\text{where } f_\stackid ~ (\underscore,\underscore,(\underscore,\underscore,\underscore,\stackid')) \triangleq \stackid \not = \stackid' \\
\end{aligned}\]

\noindent
Here, we are capturing the entire span that the system spends {\em away} from a given
coroutine.

\definition
The \emph{protected set} of a context, \(\prot_Y(\domMap,\rtm,\yts,\stackid)\),
is the set of components \(\component\) such that
\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).

\definition
The \emph{secret set} of a context, \(\secret_Y(\domMap,\rtm,\yts,\stackid)\),
is the set of components \(\component\) such that
\(\domMap ~ \component = \instack{\stackid'}{\underscore}\) where \(\stackid \not = \stackid`\).
